Cudele: A File System with Programmable Consistency and Durability

### 2017-03-28 paper/sections/3_posix_overheads.tex
paper: remove fault tolerance and add programmability section
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>
paper: cleanup terminology in section 4
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

\subsection{Durability}
\label{sec:fault-tolerance}

\begin{figure}[tb] \centering
\includegraphics[width=1\linewidth]{./figures/journal.png} 
\caption{CephFS uses a journal to stage updates and tracks dirty metadata in
the collective memory of the MDSs. Each MDS maintains its own journal, which is
broken up into 4MB segments. These segments are pushed into RADOS and deleted
when that particular segment is trimmed from the end of the log. In addition to
journal segments, RADOS also stores per-directory objects. \label{fig:journal}}
\end{figure}
\begin{figure*}[t]
  \centering
  \begin{subfigure}[b]{.3\linewidth}
      \centering
      \includegraphics[width=1.0\linewidth]{graphs/slowdown-journal.png}
      \caption{Journal Segment Size} \label{fig:slowdown-journal-a}
  \end{subfigure}
  \begin{subfigure}[b]{.3\linewidth}
      \centering
      \includegraphics[width=1.0\linewidth]{graphs/behavior-journal.png}
      \caption{180MB Segment vs. No Journal}
      \label{fig:slowdown-journal-b}
  \end{subfigure}
  \begin{subfigure}[b]{.3\linewidth}
      \centering
      \includegraphics[width=1.0\linewidth]{graphs/slowdown-cache.png}
      \caption{Inode Cache}
      \label{fig:slowdown-journal-c}
  \end{subfigure}
  \caption{The overhead of the file system metadata journal. The segment size
  is the threshold that the metadata server starts
  trimming.\label{fig:slowdown-journal}}
\end{figure*}

>>>>

\subsection{Durability}
\label{sec:durability}

\begin{figure}[tb] \centering
\includegraphics[width=1\linewidth]{./figures/journal.png} 
\caption{CephFS uses a journal to stage updates and tracks dirty metadata in
the collective memory of the metadata servers. Each metadata server maintains its own journal, which is
broken up into 4MB segments. These segments are pushed into RADOS and deleted
when that particular segment is trimmed from the end of the log. In addition to
journal segments, RADOS also stores per-directory objects. \label{fig:journal}}
\end{figure}
\begin{figure*}[t]
  \centering
  \begin{subfigure}[b]{.3\linewidth}
      \centering
      \includegraphics[width=1.0\linewidth]{graphs/slowdown-journal.png}
      \caption{Journal Segment Size} \label{fig:slowdown-journal-a}
  \end{subfigure}
  \begin{subfigure}[b]{.3\linewidth}
      \centering
      \includegraphics[width=1.0\linewidth]{graphs/behavior-journal.png}
      \caption{180MB Segment vs. No Journal}
      \label{fig:slowdown-journal-b}
  \end{subfigure}
  \begin{subfigure}[b]{.3\linewidth}
      \centering
      \includegraphics[width=1.0\linewidth]{graphs/slowdown-cache.png}
      \caption{Inode Cache}
      \label{fig:slowdown-journal-c}
  \end{subfigure}
  \caption{The overhead of the file system metadata journal. The segment size
  is the threshold that the metadata server starts
  trimming.\label{fig:slowdown-journal}}
\end{figure*}

### 2017-03-28 paper/sections/4_methodology.tex
paper: cleanup terminology in section 4
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

% describe the figure
Figure~\ref{fig:decouple} shows the mechanisms (labeled arrows) in Cudele and
which entity they are performed by (gray boxes). The metadata store and journal
are different ways of representing the namespace. The metadata store represents
the namespace as a tree of directory fragments and is easier to read/traverse.
On the other hand, the journal represent the namespace as a list of events. It
a ``pile system"; writes are fast but reads are slow because state must be
reconstructed.  Specifically, reads are slow because there is more data to
read, it is unorganized, and many of the updates may be redundant.

% describe the mechanisms
Cudele presents 6 mechanisms: RPCs, Stream, Create, Volatile Apply, Save, and
Persist. ``RPCs" does round trip remote procedure calls to establish
consistency; it is the default implementation for complying with POSIX in
CephFS. ``Stream" has the metadata servers stream a journal of metadata updates
into the object store. ``Create" allows clients to append metadata events to an
in-memory journal. ``Volatile apply" takes the in-memory journal on the client
and applies it directly to the in-memory metadata store of the metadata server
cluster. ``Save" takes the in-memory journal and writes it to the client's
disk. ``Persist" saves the journal as a an object in the object store from the
client.

@@

\begin{table}[t]
\begin{center}
\caption{Future programmers can explore the consistency (C) and
durability (D) spectrums by composing Cudele mechanisms. The consistency
and durability properties are not guaranteed until all mechanisms in the cell
are complete (i.e. the compositions should be considered atomic) and there are
no guarantees while tranisitioning between policies. \label{table:spectrum}}
\begin{tabular}{ l | l | l | l }
  C \(\rightarrow\) &&& \\  
  D \(\downarrow\)  & invisible         & eventual        & strong  \\\hline
  none                       & create            & create          & RPCs    \\
                             &                   & +volatile apply &         \\\hdashline
  local                      & create            & create          & RPCs    \\
                             & +local persist    & +local persist  & +local  \\
                             &                   & +volatile apply &  persist\\\hdashline
  global                     & create            & create          & RPCs    \\
                             & +global persist   & +global persist & +stream \\
                             &                   & +volatile apply &         \\
\end{tabular}
\end{center}
\end{table}

% describe table
The spectrum of consistency and durability guarantees that adminstrators
can construct is shown in Table~\ref{table:spectrum}. The columns are the
different consistency semantics and the rows cover the spectrum of durability
guarantees. For consistency: ``none" means the system does not handle
merging updates into a global namespace and it is assumed that middleware or
the application manages consistency lazily; ``eventual" merges updates either
when the system has time (e.g., by a background daemon) or when the client is
done writing; and updates in ``global" consistency are seen immediately by all
clients. For durability, ``none" means that updates are volatile and will
be lost on a crash of any component. Stronger guarantees are made with
``local", which means updates will be retained if the client node recovers, and
``global", where all updates are always recoverable.

% which system they represent and which are impossible
The cells in Table~\ref{table:spectrum} encompass many exisiting storage
systems. POSIX systems like CephFS and IndexFS have global consistency and
durability; DeltaFS has consistency and durability set to ``none";
and BatchFS uses ``eventual" consistency and ``local" durability. These
are just a few of the HPC examples.

@@

% what is impossible
Although we can achieve all permutations of the different guarantees in
Table~\ref{table:spectrum}, not all of them make much sense. For example, it
makes little sense to do \texttt{creates+RPCs} since both steps do the same
thing or \texttt{stream+save} since global durability is stronger and has
more overhead than local durability.  Formulaically, valid steps
include:\\

@@

\subsection{Cudele Namespace API:\\Per-Subtree Policies}

% file type
To assign consistency and durability to the subtrees we store policies in
the directory inode. This approach uses the File Type interface from the
Malacology programmable store system~\cite{} and it tells clients how to access
the underlying data or metadata. The underlying implementation stores
executable code in the inode that calls the different Cudele mechanisms. Of
course, there are many security and access control aspects of this approach but
that is beyond the scope of this paper.

% the interface
The interface for setting the subtree policies is with \{path,
<block|overwrite>, pre-allocated inodes\} tuples. For example:

@@

Given these default values decoupling the namespace with an empty policies file
would give the application 100 inodes but the subtree would behave like the
existing CephFS implementation. Table~\ref{} shows how one would use the Cudele
API to implement policies from related work.

@@

For \texttt{block}, any requests to this part of the namespace return with
``Device is busy", which will spare the MDS from wasting resources for updates
that may get overwritten. If the application does not mind losing updates, for
example it wants approximations for results that take too long to compute, it
can select \texttt{overwrite}. In this case, metadata will be written and the
computation from the decoupled namespace will take priority because the resutls
are more accurate.

@@

% what is programmability
A programmable storage system exposes internal subsystem as building blocks for
higher level services. This `dirty-slate' approach limits redundant code and
leverages the robustness of the underlying storage system. Cudele uses this
approach to great success and requires only:

@@

>>>>

% describe the figure
Figure~\ref{fig:decouple} shows the mechanisms (labeled arrows) in Cudele and
which entity they are performed by (gray boxes). The metadata store and journal
are different ways of representing the namespace. The metadata store represents
the namespace as a tree of directory fragments and is easier to read/traverse.
The journal represent the namespace as a list of events. It is
a ``pile system"; writes are fast but reads are slow because state must be
reconstructed.  Specifically, reads are slow because there is more data to
read, it is unorganized, and many of the updates may be redundant.

% describe the mechanisms
Cudele presents 6 mechanisms: RPCs, Stream, Create, Volatile Apply, Local Persist, and Global
Persist. ``RPCs" does round trip remote procedure calls to establish
consistency; it is the default implementation for complying with POSIX in
CephFS. ``Stream" has the metadata servers stream a journal of metadata updates
into the object store. ``Create" allows clients to append metadata events to an
in-memory journal. ``Volatile apply" takes the in-memory journal on the client
and applies it directly to the in-memory metadata store of the metadata server
cluster. ``Local Persist" takes the in-memory journal and writes it to the client's
disk. ``Global Persist" saves the journal as a an object in the object store from the
client.

@@

\begin{table}[t]
\begin{center}
\caption{Future programmers can explore the consistency (C) and
durability (D) spectrums by composing Cudele mechanisms. The consistency
and durability properties are not guaranteed until all mechanisms in the cell
are complete ({\it i.e.} the compositions should be considered atomic) and there are
no guarantees while tranisitioning between policies. \label{table:spectrum}}
\begin{tabular}{ l | l | l | l }
  C \(\rightarrow\) &&& \\  
  D \(\downarrow\)  	     & invisible         & eventual        & strong  \\\hline
  none                       & create            & create          & RPCs    \\
                             &                   & +volatile apply &         \\\hdashline
  local                      & create            & create          & RPCs    \\
                             & +local persist    & +local persist  & +local  \\
                             &                   & +volatile apply &  persist\\\hdashline
  global                     & create            & create          & RPCs    \\
                             & +global persist   & +global persist & +stream \\
                             &                   & +volatile apply &         \\
\end{tabular}
\end{center}
\end{table}

% describe table
The spectrum of consistency and durability guarantees that adminstrators can
construct is shown in Table~\ref{table:spectrum}. The columns are the different
consistency semantics and the rows cover the spectrum of durability guarantees.
For consistency: ``invisible" means the system does not handle merging updates
into a global namespace and it is assumed that middleware or the application
manages consistency lazily; ``eventual" merges updates at some time in the
future ({\it e.g.}, when the system has time, when number of updates reaches a
certain threshold, when the client is done writing, etc.); ; and updates in
``global" consistency are seen immediately by all clients. For durability,
``none" means that updates are volatile and will be lost on a failure. Stronger
guarantees are made with ``local", which means updates will be retained if the
client node recovers, and ``global", where all updates are always recoverable.

% which system they represent and which are impossible
Existing, state-of-the-art systems in HPC can be represented by the cells in
Table~\ref{table:spectrum}.  POSIX-compliant systems like CephFS and IndexFS
have global consistency and durability; DeltaFS uses ``invisible" consistency
and ``local" durability and BatchFS uses ``eventual" consistency and ``local"
durability. These are just a few of the HPC examples.

@@

% what is impossible
Although we can achieve all permutations of the different guarantees in
Table~\ref{table:spectrum}, not all of them make sense. For example, it
makes little sense to do \texttt{creates+RPCs} since both steps do the same
thing or \texttt{stream+save} since ``global" durability is stronger and has
more overhead than ``local" durability.  Formulaically, valid steps
include:\\

@@

\subsection{Cudele Namespace API:\\Per-Subtree Policies}
\label{sec:cudele-namespace-api-per-subtree-policies}

% the interface
The interface for setting the subtree policies is with \{path,
\(<\)block\(|\)overwrite\(>\), pre-allocated inodes\} tuples. For example:

@@

Given these default values decoupling the namespace with an empty policies file
would give the application 100 inodes but the subtree would behave like the
existing CephFS implementation. Table~\ref{table:spectrum} shows how one would use the Cudele
API to implement policies from related work.

@@

For \texttt{block}, any requests to this part of the namespace returns with
``Device is busy", which will spare the metadata server from wasting resources for updates
that may get overwritten. If the application does not mind losing updates, for
example it wants approximations for results that take too long to compute, it
can select \texttt{overwrite}. In this case, metadata will be written and the
computation from the decoupled namespace will take priority at merge time because the results
are more accurate.

@@

% what is programmability
A programmable storage system exposes internal subsystem as building blocks for
higher level services. This `dirty-slate' approach limits redundant code and
leverages the robustness of the underlying storage system. Cudele uses this
approach and re-uses some of the building blocks from the Malacology
programmable storage system~\ref{sevilla:eurosys17} to great success and
requires only:

@@

% file type
To assign consistency and durability to the subtrees we store policies in
the directory inode. This approach uses the File Type interface from the
Malacology programmable store system~\cite{sevilla:eurosys17} and it tells clients how to access
the underlying data or metadata. The underlying implementation stores
executable code in the inode that calls the different Cudele mechanisms. Of
course, there are many security and access control aspects of this approach but
that is beyond the scope of this paper.
